### 2011-12-29 joint action and knowing others minds/interacting mindreaders.2.tex
first draft done incl abstract

\begin{abstract}
%Goal ascription is fundamental for mindreading.
What evidence grounds ascriptions of thoughts and actions,
and how does the evidence support the ascriptions?
In answering this question, 
philosophers sometimes focus on mere observation, ignoring interaction.
The only evidence considered is 
 evidence that would be available to mindreaders 
observing their targets but in no position to interact with them.
The present paper,
which focuses on goal ascription,
argues that this is a mistake
and
identifies evidence 
available only to mindreaders capable of interacting with their targets.
Being poised to interact with others may make it possible to know things about their thoughts and actions which one might not otherwise be in a position to know.
This has consequences for the possible roles of interaction in explaining the evolution and development of mindreading.

%What evidence grounds ascriptions of goals to other agents' actions,
%and how does the evidence support the ascriptions?
%It is usually assumed that, in answering this question, 
%we can focus on observation and ignore interaction.
%We can restrict ourselves to evidence that would be available to mindreaders who observe their targets but are not necessarily in a position to interact with them.
%The present paper identifies evidence 
%which supports goal ascriptions
%and 
%is available only to mindreaders who are capable of interacting with their targets.

@@

Even if the hunch turns out to be right, why investigate it?
One reason is that the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
Another motive concerns the emergence, in evolution and in development, of mindreading.
Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.
One view is that needs to interact with others have driven and shaped some aspects of cognition.%
\footnote{
E.g.\ \citet[p.\ 103]{Knoblich:2006bn} suggest that 
`functions traditionally considered hallmarks of individual cognition originated through the need to interact with others'
and that
`perception, action, and cognition are grounded in social interaction.'
}
We shall not speak to this view here.
Another view is that abilities to interact with others
may have fostered the emergence in evolution---or, on a related view, in development---%
of sophisticated forms of cognition including social cognition and, in particular, mindreading.%
\footnote{
\citet[p.\ 1]{Moll:2007gu}
argue for the `Vygotskian Intelligence Hypothesis' according to which `the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation.'
See also
	\citet{Hughes:2004zj},
	\citet{Hughes:2006fu},
	\citet{Tomasello:2007gl} and
	\citet{tomasello:2008origins}.
}
Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of mindreading.

@@

%*** remove this line: to avoid annoying error only
\

@@

\section{Communicative gestures}
When introducing the problem of opaque means 
(in section \vref{sec:opaque_means})
we saw that it could  affect
communicative actions.
To illustrate this suggestion we drew on 
an experiment by \citet{hare_chimpanzees_2004}
in which chimpanzees had to find a reward 
and were helped by being shown either a failed reach or  a helpful point to the target location.
Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
Taking this paradigm as a case study, 
we want to suggest that your-goal-is-my-goal might 
enable us to understand how abilities to engage in joint action 
could be part of what enables mindreaders to make the transition from 
a simple understanding of goals 
to 
an early understanding communicative actions.%
\footnote{
Note that although our discussion borrows an experimental paradigm,
our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
that your-goal-is-my-goal
could \emph{in principle} enable
individuals to make the transition 
from
 a simple understanding of goals
to
 an early understanding of communicative actions.
Of course it would powerfully demonstrate the relevance of our theoretical argument if we could provide evidence to show that this actually happens.
But for now we are concerned with more narrowly conceptual issues.
}

@@

So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking for an individual she must associate an utterance of `daddy' with the appropriate person.%
\footnote{
Such associations appear to be in place by six months of age or earlier \citep{tincoff:1999_beginnings_,tincoff:2011}.
There is debate on when infants are first able to rapidly associate words with novel objects \citep{werker:1998_acquisition,friedrich:2011_word_}.
Interestingly,
a possible developmental decalage separating 
two measures of whether infants have formed a word-object association,
preferential looking and comprehension as measured by reaching in response to a request  \citep{gurteen:2011_rapid},
may suggest that
infants can form a word-object association before they can comprehend communicative actions involving the word.
}
As we saw, 
outside the context of joint action,
merely associating a gesture with its referent falls short of being able to respond appropriately.
After all, the goals of the gesture may well be unrelated to the goals of the interpreter's actions.
But if an interpreter supposes that her target is willing to engage in joint action with her,
then she may infer that the goal of her target's action is her goal
and so be motivated to treat the thing associated with a communicative gestures as relevant to the goal of her own actions.
This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
she may be in a position to realise, further, that the same tool can be used in other contexts.

This, in barest outline, is 
how
abilities to engage in joint action
mean that 
a mindreader with an ability to ascribe simple goals only and no understanding of communicative intent
might 
nevertheless reliably respond appropriately to some communicative gestures,
and so come be in a position to understand how such gestures can be used to guide others' actions.%
\footnote{
Contrast Csibra's claim that,
early in human development,
goal ascription (`teleological understanding' in his terms)
and 
identifying the referents of communicative gestures (`referential understanding')
`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
and
are initially two distinct `action interpretation systems'   \citep[p.\ 456]{Csibra:2003kp}.
%(although of course they come together later in development)
We have not shown that this hypothesis wrong.
But we have shown that there is another possibility:
the referential stance may 
emerge from
the teleological stance
 together with abilities to engage in simple forms of joint action.
}
Of course we have only argued that this transition is theoretically possible;
we have not attempted to defend any hypothesis about 
anything's evolution or development.
Our point is just to illustrate one of several ways in which
interaction may matter for mindreading.
Widening the evidential basis for goal ascription
to include evidence available only to interpreters capable of joint action with their targets
makes possible 
novel hypotheses about the emergence, in evolution or development (and perhaps even both), of communicative gestures.

@@

\section{The problem of false belief}
\label{sec:false_belief}
So far we have been arguing for the value of interaction to interpreters 
on the grounds that 
interacting interpreters, unlike mere observers, have 
a route to knowledge of the goals of others' actions
which avoids the problem of opaque means.
There is another reason for supposing that interaction matters for interpretation.
To introduce this reason we must first describe another problem affecting interpreters, the problem of false belief.

The problem of opaque means occurs when interpreters  must rely entirely on observation and cannot identify to which ends actions are means.
A yet more familiar problem affecting goal ascription 
arises from the interdependence of beliefs and goals.
To illustrate, 
imagine sitting at a table.
On the table are two closed opaque boxes.
One box contains an owl, the other a cat.
If the goal of your action is to retrieve the cat, 
and you believe that the cat is in the north box,
then 
(unless things are going very badly) 
you will reach for the north box.
But of course if you had believed instead that the cat was in the south box,
then, in acting on the same goal, you would have reached for the south box.
Now consider Ayesha who is observing your actions.
Suppose Ayesha has sufficient reason to believe, falsely, that you know the cat is in the south box.
Then she may be justified in supposing, incorrectly, that 
the goal of your action, in reaching for the north box,
is to retrieve the owl.
As this illustrates,
differences in belief between observers and protagonists can 
impair goal ascription
when the interpreter is unaware of those differences.
Call this the problem of false belief.

@@

Exploiting your-goal-is-my-goal does not make it possible to  avoid the problem of false belief entirely,
it only shifts  the problem.
To illustrate,
in the above example Ayesha's ability to correctly identify the goal of your action
depends on your correctly anticipating the goals of her actions---on your knowing that she is concerned with retrieving the cat.
If you incorrectly anticipated that her actions would be directed to retrieving the owl,
Ayesha would have been mistaken in taking your actions to be directed to retrieving the cat.
As this indicates,
exploiting your-goal-is-my-goal 
can shift 
both 
\emph{which} differences in belief have the potential to impair  goal ascription
and also 
\emph{who} needs to be aware of those differences.

In short, then, abilities to engage in joint action provide an interpreter with a route to knowledge of the goals of other agents' actions which does not depend on her knowledge of what her target believes.
This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
Rather they shift the burden of resolving the problem of false belief from an interpreter to her target.
This is potentially valuable for interpreters who have limited insight into their targets' beliefs,
or who lack abilities to track differences in belief altogether.

On a purely observational model of interpretation,
it seems inevitable that an interpreter who is unable to track others' beliefs 
will have to proceed as if there were no differences in beliefs.
As several philosophers' arguments suggest,
it is plausible that 
the interdependence of beliefs and goals 
means that
an interpreter will always be able to assign goals to others' actions 
in a way that is consistent with their beliefs not differing from her own
(*refs: Davidson; \citealp[p.\ 49]{Bennett:1976rg}).
If so,
on a purely observational model of interpretation
there is no straightforward way in which
the need to track differences in belief 
will become apparent to an interpreter.
Things are different
if we extend our model of interpretation
to take into account the possibility of interaction.
We can then see how
an interpreter who is unable to track others' beliefs
will encounter anomalies.
For some differences in belief will mean that
 ordinary, purely observational routes to knowledge will result in ascribing one goal
 whereas
 exploiting your-goal-is-my-goal would result in a ascribing a different goal.
That is, the evidence available to an interpreter may inexplicably support two incompatible ascriptions. 
In this way,
differences in belief have the potential to cause puzzling anomalies
even before the interpreter is able to track such differences.
This 
raises the possibility
that abilities to engage in joint action
may play a role in explaining how  awareness of belief emerges.

\section{Conclusion}
Our aim was to show that interaction can facilitate mindreading in this sense:
some routes to knowledge are closed to mindreaders who rely exclusively on observation
but
open to interacting mindreaders.
In pursuing this aim we focused on interactions which are joint actions.
Our suggestion was this.
Where an interpreter recognizes that her target is willing to engage in joint action
and is unwilling to change which goals her own actions will be directed to,
she may be in a position to know
that the goals of her target's actions will be her goals.
This in outline is the `your-goal-is-my-goal' route to knowledge.

@@

To show that this route to knowledge is potentially valuable,
we argued that it enables interpreters to overcome two problems,
the problem of opaque means
and the problem of false belief.
These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
But interest in the evolution or development of mindreading
motivates focusing on 
interpreters with
  absent or fledgling communicative skills,
  limited insight into others' minds
  and narrow knowledge of how things work.
For such interpreters
the problems of false belief and opaque means may easily arise,
particularly in cases involving novel tools 
and 
communicative gestures.
Interpreters 
could overcome the two problems 
by exploiting the your-goal-is-my-goal route to knowledge.
It follows that 
expertise with 
  tools 
  and
  communicative gestures
does not presuppose knowledge of others' intentions and beliefs.
Such expertise,
far from  presupposing sophisticated forms of mindreading,
might instead
play a role in explaining their emergence in evolution or development (or both).

How might this work?
***sketch the story***
Clearly this is a wild conjecture,
and not one that we have given any reason for accepting.
Our point in mentioning it is merely to illustrate how
 better understanding the role of interaction in goal ascription and mindreading generally
may eventually lead to discoveries
on how
 sophisticated forms of cognition
 emerge from
 conceptually and cognitively undemanding forms of interaction.
If the culture of tool use,
the simpler forms of communicative gesture
or the sharing of knowledge and other mental states
originates in abilities to interact,
then we need to know how interaction could lead to the emergence of these things.
Our  suggestion has been that
understanding how interacting mindreaders 
could know things about their minds which mere observers might not be in a position to know
will provide some of the details.

>>>>

\begin{abstract}
Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
This paper defends a positive answer.
Mindreading is sometimes reciprocal: 
the mindreader's target reciprocates by taking the mindreader as a target for mindreading.
The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour 
where mindreaders are, or appear to be, in a position to interact.
It follows that
revisions and extensions 
are needed for 
standard theories of mindreading
on which
all the evidence relevant to mindreading 
is in principle available to manifestly passive mindreaders.
The view also has consequences for 
understanding 
how abilities to interact combined with comparatively simple forms of mindreading
may explain the emergence,
in evolution or development,
of sophisticated forms of social cognition.

%
%What evidence grounds ascriptions of thoughts and actions,
%and how does the evidence support the ascriptions?
%In answering this question, 
%philosophers sometimes focus on mere observation, ignoring interaction.
%The only evidence considered is 
% evidence that would be available to mindreaders 
%observing their targets but in no position to interact with them.
%The present paper,
%which focuses on goal ascription,
%argues that this is a mistake
%and
%identifies evidence 
%available only to mindreaders capable of interacting with their targets.
%Being poised to interact with others may make it possible to know things about their thoughts and actions which one might not otherwise be in a position to know.
%This has consequences for the possible roles of interaction in explaining the evolution and development of mindreading. 
%
%What evidence grounds ascriptions of goals to other agents' actions,
%and how does the evidence support the ascriptions?
%It is usually assumed that, in answering this question, 
%we can focus on observation and ignore interaction.
%We can restrict ourselves to evidence that would be available to mindreaders who observe their targets but are not necessarily in a position to interact with them.
%The present paper identifies evidence 
%which supports goal ascriptions
%and 
%is available only to mindreaders who are capable of interacting with their targets.

@@

Even if the hunch turns out to be right, why investigate it?
One reason is that the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
Another motive concerns the emergence, in evolution and in development, of mindreading.
Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.
One view is that needs to interact with others have driven and shaped some aspects of cognition.%
\footnote{
E.g.\ \citet[p.\ 103]{Knoblich:2006bn} suggest that 
`functions traditionally considered hallmarks of individual cognition originated through the need to interact with others'
and that
`perception, action, and cognition are grounded in social interaction.'
}
We shall not speak to this view here.
Another view is that abilities to interact with others
may have fostered the emergence in evolution---or, on a related view, in development---%
of sophisticated forms of cognition including social cognition and, in particular, mindreading.%
\footnote{
\citet[p.\ 1]{Moll:2007gu}
argue for the `Vygotskian Intelligence Hypothesis' according to which `the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation.'
See also
	\citet{Hughes:2004zj},
	\citet{Hughes:2006fu},
	\citet{Tomasello:2007gl} and
	\citet{tomasello:2008origins}.
}
Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of sophisticated forms of mindreading.

@@

@@

\section{Communicative gestures}
\label{sec:communicative_gestures}
When introducing the problem of opaque means 
(in section \vref{sec:opaque_means})
we saw that it could  affect
communicative actions.
To illustrate this suggestion we drew on 
an experiment by \citet{hare_chimpanzees_2004}
in which chimpanzees had to find a reward 
and were helped by being shown either a failed reach or  a helpful point to the target location.
Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
Taking this paradigm as a case study, 
we want to suggest that your-goal-is-my-goal might 
enable us to understand how abilities to engage in joint action 
could be part of what enables mindreaders to make the transition from 
a simple understanding of goals 
to 
an early understanding communicative actions.%
\footnote{
Note that although our discussion borrows an experimental paradigm,
our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
that your-goal-is-my-goal
could \emph{in principle} enable
individuals to make the transition 
from
 a simple understanding of goals
to
 an early understanding of communicative actions.
Of course it would powerfully demonstrate the relevance of our theoretical argument if we could provide evidence to show that this actually happens.
But for now we are concerned with more narrowly conceptual issues.
}

@@

So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking to see who has an object she must associate an utterance of `daddy' with the daddy.%
\footnote{
Such associations appear to be in place by six months of age or earlier \citep{tincoff:1999_beginnings_,tincoff:2011}.
There is debate on when infants are first able to rapidly associate words with novel objects \citep{werker:1998_acquisition,friedrich:2011_word_}.
Interestingly,
a possible developmental decalage separating 
two measures of whether infants have formed a word-object association,
preferential looking and comprehension as measured by reaching in response to a request  \citep{gurteen:2011_rapid},
may suggest that
infants can form a word-object association before they can comprehend communicative actions involving the word.
}
As we saw, 
outside the context of joint action,
merely associating a gesture with its referent falls short of being able to respond appropriately.
But if an interpreter supposes that her target is willing to engage in joint action with her,
then she may infer that the goal of her target's action is her goal
and so be motivated to treat the thing associated with a communicative gesture as relevant to the goal of her own actions.
This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
she may be in a position to realise, further, that the same tool can be used in other contexts.

This, in barest outline, is 
how
possessing abilities to engage in joint action
mean that 
an individual with an ability to ascribe simple goals only and no understanding of communicative intent
might 
nevertheless reliably respond appropriately to some communicative gestures,
and so come be in a position to understand how such gestures can be used to guide others' actions.%
\footnote{
Contrast Csibra's claim that,
early in human development,
goal ascription (`teleological understanding' in his terms)
and 
identifying the referents of communicative gestures (`referential understanding')
`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
and
are initially two distinct `action interpretation systems'   \citep[p.\ 456]{Csibra:2003kp}.
%(although of course they come together later in development)
We have not shown that this claim wrong.
But we have shown that there is another possibility:
the referential stance may 
emerge from
the teleological stance
 together with abilities to engage in simple forms of joint action.
}
Of course we have only argued that this transition is theoretically possible;
we have not attempted to defend any hypothesis about 
anything's evolution or development.
Our point is just to illustrate one of several ways in which
interaction may matter for mindreading.
Widening the evidential basis for goal ascription
to include evidence available only to mindreaders actually or apparently capable of joint action with their targets
makes possible 
novel hypotheses about the emergence, in evolution or development (and perhaps both), of communicative gestures.

@@

\section{The problem of false belief}
\label{sec:false_belief}
So far we have been arguing for the value of interaction for goal ascription
on the grounds that 
interacting mindreaders, unlike mere observers, have 
a route to knowledge of the goals of others' actions
which avoids the problem of opaque means.
There is another reason for supposing that interaction matters for goal ascription and (at least indirectly) for mindreading more generally.
To introduce this reason we must first describe another problem affecting goal ascription, the problem of false belief.

Recall that the problem of opaque means occurs when interpreters  must rely entirely on observation and cannot identify to which ends actions are means 
(see section \vref{sec:opaque_means}).
A yet more familiar problem affecting goal ascription 
arises from the interdependence of beliefs and goals.
To illustrate, 
imagine sitting at a table.
On the table are two closed opaque boxes.
One box contains an owl, the other a cat.
If the goal of your action is to retrieve the cat, 
and you believe that the cat is in the north box,
then 
(unless things are going very badly) 
you will reach for the north box.
But of course if you had believed instead that the cat was in the south box,
then, in acting on the same goal, you would have reached for the south box.
Now consider Ayesha who is observing your actions.
Suppose Ayesha has sufficient reason to believe, falsely, that you know the cat is in the south box.
Then she may be justified in supposing, incorrectly, that 
the goal of your action, in reaching for the north box,
is to retrieve the owl.
As this illustrates,
differences in belief between observers and protagonists can 
impair goal ascription
when the interpreter is unaware of those differences.
Call this the problem of false belief.

@@

Exploiting your-goal-is-my-goal does not make it possible to  avoid the problem of false belief entirely,
it only shifts  the problem.
To illustrate,
in the above example Ayesha's ability to correctly identify the goal of your action
depends on your correctly anticipating the goals of her actions---on your knowing that she is concerned with retrieving the cat.
If you incorrectly anticipated that her actions would be directed to retrieving the owl,
Ayesha would have been mistaken in taking your actions to be directed to retrieving the cat.
As this indicates,
exploiting your-goal-is-my-goal 
can shift 
both 
\emph{which} differences in belief have the potential to impair  goal ascription
and also 
\emph{who} needs to be aware of those differences.
This is a second illustration of how the possibility of interaction 
may enable 
	even individuals with only limited insights into others' minds and actions
 	to exploit reciprocity in mindreading---to exploit the fact that their targets are sometimes themselves mindreading the mindreaders.

In short, then, abilities to engage in joint action open a route to knowledge of the goals of other agents' actions which does not depend on the goal ascriber knowing what her target believes even when their beliefs relevantly differ.
This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
Rather they shift the burden of resolving the problem of false belief from an interpreter to her target.
This is potentially valuable for interpreters who have limited insight into their targets' beliefs,
or who lack abilities to track differences in belief altogether.

On a purely observational model of mindreading,
it seems inevitable,
given that observed behaviours will be consistent with many different ascriptions of beliefs and goals,
that a goal ascriber who was unable to track others' beliefs 
would often have to proceed as if there were no relevant differences in beliefs.%
\footnote{
Compare \citet[p.\ 49]{Bennett:1976rg}.  For a related argument concerning beliefs and desires rather than beliefs and goals, see 
	\citet[p.\ 17]{Davidson:1984pr}
%what can fairly be treated as evidence for any one of the propositional attitudes can count as evidence only if untested assumptions are made about others.”  [1984d, p. 17]
and
	(\citeyear[pp.\ 159--167]{Davidson:1984wh}).
%`This creates a problem, for it means that behaviour, which is the main evidential basis for attributions of belief and desire, is reckoned the result of two forces less open to public observation. Thus where one constellation of beliefs and desires will rationalize an action, it is always possible to find a quite different constellation that will do as well. Even a generous sample of actions threatens to leave open an unacceptably large number of alternative explanations.'
%`of course it cannot be assumed that speakers never have false beliefs. Error is what gives belief its point. We can, however, take it as given that most beliefs are correct. The reason for this is that a belief is identified by its location in a pattern of beliefs; it is this pattern that determines the subject matter of the belief, what the belief is about. Before some object in, or aspect of, the world can become part of the subject matter of a belief (true of false) there must be endless true beliefs about the subject matter. False beliefs tend to undermine the identification of the subject matter; to undermine, therefore, the validity of a description of the belief as being about that subject.'
}
So on a purely observational model of interpretation
there may fail to be any straightforward way in which
the need to track differences in belief 
could become apparent to a mindreader who did not already understand that beliefs can be false.
Things are different
if we extend our model of mindreading
to take into account the possibility of interaction.
We can then see how
a goal ascriber who is unable to track others' beliefs
will encounter anomalies.
For some differences in belief 
(the mere fact of such differences, not anyone's awareness of them)
will mean that
 ordinary, purely observational routes to knowledge will result in ascribing one goal
 whereas
 exploiting your-goal-is-my-goal would result in a ascribing a different goal.
That is, the evidence available to a goal ascriber may inexplicably support two incompatible ascriptions. 
In this way,
differences in belief have the potential to cause puzzling anomalies
even before a mindreader is able to track such differences.
This 
raises the possibility
that abilities to engage in joint action
may play a role in explaining how  awareness of belief emerges.%
\footnote{
Some researchers have suggested that awareness of belief is innate in humans
\citep{Leslie:1998nq,Leslie:2005ef,Baillargeon:gx},
although this view is currently controversial
\citep{Apperly:2009ju,Ruffman:2005yl}.
But whether or not the emergence in development of awareness of belief stands in need of explanation,
few would doubt that its emergence in evolution does need explaining.
}

\section{Conclusion}
Our aim was to show that interaction can facilitate mindreading in this sense:
some routes to knowledge are closed to mindreaders who rely exclusively on observation
but
open to interacting mindreaders.
In pursuing this aim we focused on interactions which are joint actions.
Our suggestion was this.
Where a mindreader recognizes that her target is willing to engage in joint action
and is unwilling to change which goals her own actions will be directed to,
she may be in a position to know
that the goals of her target's actions will be her goals.
This in outline is the `your-goal-is-my-goal' route to  knowledge (see further section \vref{sec:your_goal_is_my_goal}).

@@

To show that this route to knowledge is potentially valuable,
we argued that it enables mindreaders to overcome two problems,
the problem of opaque means
and the problem of false belief.
These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
But interest in the evolution or development of mindreading
motivates focusing on 
those with
  absent or fledgling communicative skills,
  limited insight into others' minds
  and narrow knowledge of how things work.
For such individuals
the problems of opaque means and false beliefs may easily arise,
particularly in cases involving novel tools 
and 
communicative gestures.
Both problems can be overcome 
by exploiting the your-goal-is-my-goal route to knowledge.
It follows that 
expertise with 
  tools 
  and
  communicative gestures
does not presuppose knowledge of others' intentions and beliefs.
This shows that it is coherent to suppose that 
 expertise with tools or communicative gestures (or both),
far from  presupposing sophisticated forms of mindreading,
might instead
play a role in explaining their emergence in evolution or development.

How might this work?
Here is a wild conjecture about
a path from simple forms of interaction to sophisticated forms of mindreading.
We have seen that 
	abilities to ascribe goals to actions 
 together with
	abilities to engage in simple forms of joint action involving distributive goals
 may be sufficient to explain how individuals come to productively misunderstand communicative gestures in the context of joint action (see section \vref{sec:communicative_gestures}).
This might  be a step towards responding appropriately to, and using, communicative gestures in other contexts,
which might then play a role in explaining how early forms of communication emerge.
But we know that early, non-linguistic forms of communication are essential for the later appearance of linguistic communication,
and abilities to communicate by language in turn appear to play an important role in the development (at least) of sophisticated forms of mindreading.%
\footnote{
A range of evidence is reviewed in contributions to \citet{Astington2005ot}.
}
Now clearly this wild conjecture is not one that we have provided any reason for accepting.
Our point in mentioning it is merely to illustrate how
 better understanding the role of interaction in goal ascription and  mindreading more generally
might eventually lead to discoveries
on how
 sophisticated forms of cognition
 emerge from
 conceptually and cognitively undemanding forms of interaction.
A culture of tool use,
	the flexible use of communicative gestures,
	or insight into interpersonal differences in belief and other mental states:
if any or all of these 
originate in abilities to interact,
then we need to know how interaction could lead to their emergence.
And one small step towards acquiring this knowledge might just be 
understanding how interacting mindreaders 
could 
	come to know things 
	and
	encounter anomalies
which they might not be able to know or encounter if they were mere observers.
% matters because it enables us to better understand the evidential basis of mindreading---%
%and it might just also matter because it plays a role in demonstrating the theoretical coherence of hypotheses about mindreading's emergence.
